Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities

### 2018-06-09 Conference/IRC 2018/IRC18.tex
added Conference folder

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Autonomous Quadrotor 3D Mapping and Exploration Using Exact Probabilities and Entropies}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

%\author{Evan Kaufman, Taeyoung Lee, and Zhuming Ai%, and Ira S. Moskowitz%
%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
%\thanks{Zhuming Ai, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
%\thanks{This research has been supported by the U.S. Naval Research Laboratory Base Program ``Intelligent Microflyer'' and in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
%}

%\author{\IEEEauthorblockN{Evan Kaufman, Kuya Takami, Taeyoung Lee}
%\IEEEauthorblockA{Department of Mechanical and Aerospace Engineering\\The George Washington University\\
%Washington, DC 20052\\
%Email: \{evankaufman,kuya,tylee\}@gwu.edu}
%\and
%\IEEEauthorblockN{Zhuming Ai}
%\IEEEauthorblockA{Information Management \& Decision Architectures
%              \\ U.S. Naval Research Laboratory \\
%              Washington, DC 20375}
%}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

@@

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Robotic mapping and exploration has received significant attention recently because exploring uncertain environments can pose danger to humans, and robots can produce highly detailed maps. Commonly uncertain environments contain complex geometries and robots frequently face restrictive terrains, motivating a 3D representation of mapping from a vehicle capable of hover. This paper focusses on the popular occupancy grid mapping representation, with the goal of determining occupancy probabilities of evenly-spaced grid cells in 3D. With recent developments from the authors using the stochastic properties of depth sensors directly with a rigorous, probabilistic formulation, this paper extends this precise mapping technique to 3D for real-time implementation with realistic sensor capabilities. Furthermore, important aspects of the 3D map are projected onto 2D maps, where a predicted level of map uncertainty, known as Shannon's entropy, provides an exploration policy that governs robotic motion. Both mapping and exploration algorithms produce important information that impacts each other and are demonstrated simultaneously with numerical simulations and experiments in real-time.
\end{abstract}

@@

Robotic mapping is the process of generating maps, which represent the environment surrounding a robot. 
Effectively modeling this environment is crucial when building an understanding of an initially uncertain environment.
Robotic mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge of the environment are computationally tractable.

Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment. In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}. Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which provides powerful open-source software tools for representing free and occupied spaces nearby the robot, with the capability of extending the mapping limits and changing cell sizes. However, Octomaps produce cell occupancy probabilities that do not directly depend on the depth sensor stochastic properties, known as the forward sensor model. Ideally, the probability density provided by the forward sensor model should be used explicitly when determining occupancy probabilities of grid cells conditioned on the robot pose and measurements, known as the inverse sensor model. Instead, these approaches mentioned above determine occupancy probabilities using an approximate function to replace the true inverse sensor model. Additionally, this approach replaces probabilities with log-odds ratios, then combines past and current information with a potentially harmful Markov assumption.

Alternatively, another popular approach is to simulate maps, robot poses, and measurements and use a learning algorithm, to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}, which is implemented in 3D with~\cite{SouMaiGon12}. These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

The approximate nature of the inverse sensor model motivated the authors of this paper to design computationally-efficient algorithm to construct the exact inverse sensor model that follows a Bayesian framework~\cite{KauLeeAiMos16,KauTakAiLee17}. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancies of the other cells blocked by it are irrelevant to the forward sensor model, and this model is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. This method combines a priori probabilities of occupancy and multiple range measurements according to a Bayesian framework to obtain more accurate maps. As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption or particle-filtering methods such as~\cite{TanThoWolBus14} that require particle weighting and resampling to obtain a variation on a Bayesian solution. 
This paper proposes extending this algorithm to 3D with realistic sensor limitations. We demonstrate the efficacy of the proposed algorithm using two different depth sensors simultaneously to generate a 3D occupancy grid map with accurate probabilities in real-time with numerical simulations and experiments.

The second topic of this paper deals with autonomous exploration. A common approach to solve the autonomous exploration problem is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98}. This is the process of executing actions that move the robot toward the closest boundary between visited and unvisited space, known as a frontier.
Then the robot takes measurements at this location such that the mapped territory expands, and thus the new frontiers are pushed back. This process is repeated until the map is well-known.
This frontier-based approach is extended to 3D with~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation, and similarly in~\cite{SawKriSri09} using a visibility metric.
Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map, though these systematic actions are not based on any consideration of the future uncertainty of the probabilistic map or optimality.

Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probabilities. However, the expected entropy of an action is not directly calculated. Instead, this approach assumes that expected entropy is equivalent to entropy based on expected measurement scans; a Rao-Blackwellized particle filter estimates the robot pose and map, where expected measurements are drawn from each particle when finding the expected information gains, then combined with the weighting on each particle. However, assuming an expected measurement is sufficient to predict expected information is subject to inaccuracies due to the nonlinear nature of object detection probability and entropy. For these reasons, the authors have proposed a direct calculation of expected information gain from occupancy probabilities and stochastic sensor models~\cite{KauAiLee16,KauTakAiLee17}. In this paper, this concept is extended to 3D, where grid cells above and below a predetermined exploration height are projected onto 2D maps for entropy and collision calculations. In this sense, the proposed approach is entropy-based in 3D similar to~\cite{JohStaPfaBur07} where uncertainties are projected onto a 2D map, except that the entropy measures are based on map probabilities and sensor properties directly.

The paper is organized as follows. The problem is formulated in Section II. Special considerations for mapping and exploration in 3D are established in Section III. A numerical example is demonstrated in Section IV. Experimental results are shown in Section V, followed by conclusions.

@@

The key differences between the approaches presented in this paper and others is how map cell occupancy probability and measurement ray entropy are calculated directly from a priori cell probabilities and stochastic properties of the sensors involved. These are defined next.

@@

% {R ? R3×3 | RT R = I, det R = 1}
Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray. Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ such that $x_t\in\Re^3$ and $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, where $X_t$ is assumed to be known. Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$. To obtain this probability for all cells,~\cite{KauLeeAiMos16,KauTakAiLee17} introduced a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$, and $\mathbf{r}_{k+}$ refers to the event that the $k$-th cell is the closest occupied space. Then, the inverse sensor model for the $k$-th cell is
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})\nonumber\\
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. The normalizer $\eta$ is given by
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$. Since the terms of \refeqn{Unnormalized} are easily obtained and several are used repeatedly to obtain \refeqn{allEta}, the computational cost of \refeqn{RayISMAnswer} is trivial and can be applied in real-time. This process is repeated for all measurements composing scan $Z_t$.

@@

Let an unvisited candidate future pose and its measurement scan be $X_c$ and $Z_c$, respectively. Since the robot has not visited $X_c$, any change to the probabilistic map must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Since all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these are removed from the remaining equations for simplicity. For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where subscript $k$ refers to the event that the candidate \emph{measurement} captures the $k$-th cell along the ray, i.e., $z_{c,k}$ is the distance from $x_c$ to the $k$-th grid cell. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{allEta} and entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap}. The second term is found using \refeqn{allEta} with
\begin{align}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}}.
\end{align}

The goal of entropy-based autonomous exploration is to select the optimal pose $X^*_c$ that maximizes the expected information gain,
\begin{align}
\label{eqn:objectiveFunOptimized}
X_c^*&=\argmax_{X_c}{\ \mathcal I(X_c)},
\\
\label{eqn:objectiveFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_c,Z_c))\right],
\end{align}
where \refeqn{DiscExpEntropyRay} is used repeatedly for several sample rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.

@@

The inverse sensor model from~\cite{KauLeeAiMos16,KauTakAiLee17} is based on an arbitrary vector spanning 2D space. The distances from the robot to grid cells are obtained geometrically through ray casting. This method is easily extended to 3D space, and the inverse sensor model is applied identically.

>>>>

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities}

@@

\maketitle

\begin{abstract}
%Robotic mapping and exploration have received significant attention recently because exploring uncertain environments can pose danger to humans, and robots can produce highly detailed maps. Commonly uncertain environments contain complex geometries and robots frequently face restrictive terrains, motivating a 3D representation of mapping from a vehicle capable of hover.

This paper deals with the aerial exploration for an unknown three-dimensional environment, where Bayesian probabilistic mapping is integrated with a stochastic motion planning scheme to minimize the map uncertainties in an optimal fashion. We utilize the popular occupancy grid mapping representation, with the goal of determining occupancy probabilities of evenly-spaced grid cells in 3D with sensor fusion from multiple depth sensors with realistic sensor capabilities. The 3D exploration problem is decomposed into 3D mapping and 2D motion planning for efficient real-time implementation. This is achieved by projecting important aspects of the 3D map onto 2D maps, where a predicted level of map uncertainty, known as Shannon's entropy, provides an exploration policy that governs robotic motion. Both mapping and exploration algorithms are demonstrated with both numerical simulations and quadrotor flight experiments.
\end{abstract}

@@

Robotic mapping is the process of generating maps representing the environment surrounding a robot.
Mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment and are computationally tractable.

Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment. In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}. Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which has added features such as extending the mapping limits and changing cell sizes.

However, these approaches are based on an approximate solution of the inverse sensor model, to determine the occupancy probabilities based on the depth sensor measurements. As such, the prior belief of the environment and the stochastic properties of the sensor are not properly accounted. Another popular approach for the inverse sensor model involves a simulated learning technique~\cite{Thr01,ThrBurFox05,SouMaiGon12}, but this is not suitable for the higher dimensional configurations composed of maps, poses, and measurement due to the large number of required samples. None of these approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.

%        1.      sensor fusion in mapping
The approximate nature of the occupancy probability motivates an exact Bayesian solution that can be applied in 3D. This paper proposes a computationally-efficient algorithm to obtain the exact solution to the inverse sensor problem to generate probabilities in a 3D environment, following the framework of~\cite{KauLeeAiMos16,KauTakAiLee17}. The proposed approach integrates all of the stochastic information available for the environment, such as the prior map estimate and multiple sensor measurements, according to the Bayesian framework, to obtain the exact 3D occupancy probabilities. As such, it provides a unified framework for sensor fusion between multiple sensors with various properties, such as sensor accuracy, and fields of view.

%        2.      proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration. Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
The second topic of this paper deals with autonomous exploration, which provides the policies that govern robotic motion. A common approach is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D  using the Octomap representation in~\cite{ZhuDinLinWu15,SenWan16,KleDor13}, and similarly using a visibility metric in~\cite{SawKriSri09}. This approach moves robots toward the boarder between certain and uncertain spaces, though these systematic actions are not based on any consideration of the future probabilistic map uncertainty.

Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, and it is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17}. This paper focusses on extending these concepts to 3D in a numerically efficient manner.  The prescribed 3D exact inverse sensor model is critical in computing the expected information gain precisely. There are several challenges with obtaining expected information gain of 3D measurement scans, particularly due to the extended computational load. Considering the typically low variation in floor plans along the vertical direction, this paper proposes a numerically efficient scheme that projects stochastic properties of the 3D probabilistic occupancy grid to 2D spaces for optimal information gain and collision avoidance. This exploits the important characteristics of 3D maps. This process of combining 3D mapping and 2D exploration is completed in a computationally-efficient manner.

%        3.      hardware/software for aerial exploration experiment.
Finally, we present both numerical simulations and flight experiments for the proposed exploration schemes. The multiple tasks of sensor data acquisition, mapping, exploration, flight control, and communications are implemented with Python/C++, to be executed simultaneously in a ROS environment. They are verified with a quadrotor unmanned aerial vehicle operated by the NVidia Jetson computing module with an ASUS Xtion infrared depth sensor and Hokuyo laser scanner.

The paper is organized as follows. The problem is formulated in Section II. Special considerations for mapping and exploration in 3D are established in Section III. A numerical example is demonstrated in Section IV. Experimental results are shown in Section V, followed by conclusions and future work.

%% Prior to TLee edits:
%Robotic mapping is the process of generating maps representing the environment surrounding a robot.
%Mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
%Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment and are computationally tractable. 
%
%Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment. In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}. Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which has added features such as extending the mapping limits and changing cell sizes. However, Octomaps, and the 3D uniform grid-based approaches mentioned above, produce cell occupancy probabilities that do not directly depend on the depth sensor stochastic properties. Instead, these approaches determine occupancy probabilities using an approximate function. Another method introduced in~\cite{Thr01,ThrBurFox05} and implemented in 3D with~\cite{SouMaiGon12} uses a simulated learning technique to find these probabilities, which may be difficult over the large dimensional space composed of maps, poses, and measurements. An additional approach such as~\cite{TanThoWolBus14} applies a particle filter to estimate the map. None of these approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.
%
%%        1.      sensor fusion in mapping
%The approximate nature of the occupancy probability motivates an exact Bayesian solution that can be applied in 3D. This paper proposes a computationally-efficient algorithm that follows the framework of~\cite{KauLeeAiMos16,KauTakAiLee17} to generate probabilities in a 3D environment. Since this approach only requires the origin and direction of sensor measurements to be known, the algorithm is designed to work for any number of sensors. Sensor fusion from several sources can provide important information about the environment with different levels of sensor certainty, fields of view, and vantage points.
%
%%        2.      proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration. Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
%The second topic of this paper deals with autonomous exploration, which provide the policies that govern robotic motion. A common approach is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D with~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation, and similarly in~\cite{SawKriSri09} using a visibility metric. This approach moves robots toward the boarder between certain and uncertain space, though these systematic actions are not based on any consideration of the future probabilistic map uncertainty. Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probabilities. Shannon's entropy is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17}, and this paper focusses on extending these concepts to 3D. There are several challenges with obtaining expected information gain of 3D measurement scans, particularly as they relate to computational load. Considering the typically low variation in floor plans that robots explore during indoor explorations, this paper proposes two methods to generate 2D exploration maps that exploit the important characteristics of 3D maps. This process of combining 3D mapping and 2D exploration is completed in a computationally-efficient manner.
%
%%        3.      hardware/software for aerial exploration experiment.
%Finally, this paper comments on the hardware and software necessary to implement these challenging algorithms for an aerial mapping and exploration experiment. This paper provides the basic software structure and features for mapping and generating exploration trajectories on a host computer while running depth sensors and a flight controller onboard a mobile robot.
%
%The paper is organized as follows. The problem is formulated in Section II. Special considerations for mapping and exploration in 3D are established in Section III. A numerical example is demonstrated in Section IV. Experimental results are shown in Section V, followed by conclusions and future work.

@@

The key differences between the approaches presented in this paper and others relate to how map cell occupancy probabilities and measurement ray entropies are calculated directly from a priori cell probabilities and stochastic properties of the sensors involved. These are defined next.

@@

% {R ? R3×3 | RT R = I, det R = 1}
Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray. Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ with location $x_t\in\Re^3$ and attitude $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, where $X_t$ is assumed to be known. Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$. To obtain this probability for all cells,~\cite{KauLeeAiMos16,KauTakAiLee17} introduced a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$, and $\mathbf{r}_{k+}$ refers to the event that the $k$-th cell of map $r$ is the closest occupied space. Then, the inverse sensor model for the $k$-th cell is
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})\nonumber\\
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. The normalizer $\eta$ is given by
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$. Since the terms of \refeqn{Unnormalized} are easily obtained and several are used repeatedly to obtain \refeqn{allEta}, the computational cost of \refeqn{RayISMAnswer} is trivial and can be applied in real-time. This process is repeated for all measurements composing scan $Z_t$.

@@

Let an unvisited candidate future pose and its measurement scan be $X_c$ and $Z_c$, respectively. Since the robot has not visited $X_c$, any change to the probabilistic map must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Since all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these are removed from the remaining equations for simplicity. For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where subscript $k$ refers to the event that the candidate \emph{measurement} captures the $k$-th cell along the ray, i.e., $z_{c,k}$ is the distance from $x_c$ to the $k$-th grid cell. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{allEta} and entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap}. The second term is found using \refeqn{allEta} with
\begin{align}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer calculated with $z_{c,k}$.
The goal of entropy-based autonomous exploration is to select the optimal pose $X^*_c$ that maximizes the expected information gain,
\begin{align}
\label{eqn:objectiveFunOptimized}
X_c^*&=\argmax_{X_c}{\ \mathcal I(X_c)},
\\
\label{eqn:objectiveFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_c,Z_c))\right],
\end{align}
where \refeqn{DiscExpEntropyRay} is used repeatedly for several sample rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.

@@

The inverse sensor model from~\cite{KauLeeAiMos16,KauTakAiLee17} is based on an arbitrary vector spanning 2D space. The distances from the robot to grid cells are obtained geometrically through ray casting. This method is easily extended to 3D space, and the inverse sensor model is applied identically. Since the map is updated ray-by-ray (each measurement ray is considered once by itself), this approach is extended for measurement fusion of any number of sensors; the map is updated by each ray of a scan sequentially, then the map is updated by the rays of the next scan in queue, which might have different characteristics. There are several reasons why a robot might be equipped with multiple sensors, e.g., different fields-of-view or depth sensor properties. The proposed 3D mapping approach can handle measurements from different sources with different properties.
